---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "# ",
  fig.path = "figures/README-",
  dpi = 250
)
```


# __invgamma__

__invgamma__ implements the `(d/p/q/r)` statistics functions for the [chi distribution](https://en.wikipedia.org/wiki/Chi_distribution) in [R](http://cran.r-project.org).  It is ideal for using in other packages since it is lightweight and leverages the `(d/p/q/r)chisq()` line of functions maintained by CRAN.



### Getting __chi__

<!-- There are two ways to get __chi__.  For the [CRAN version](https://cran.r-project.org/package=chi), use -->
<!-- ```{r, eval=FALSE} -->
<!-- install.packages("chi") -->
<!-- ``` -->
For the development version, use
```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("dkahle/chi")
```



### The `(d/p/q/r)chi()` functions

The defining property of the chi distribution is that it is the square root of a chi random variable.


The [PDF](https://en.wikipedia.org/wiki/Probability_density_function) (the _f(x)_) can be evaluated with the `dchi()` function:
```{r, fig.height = 3}
library(chi)
library(ggplot2); theme_set(theme_bw())
x <- seq(0, 6, .01)
qplot(x, dchi(x, 7), geom = "line")
```


The [CDF](https://en.wikipedia.org/wiki/Cumulative_distribution_function) can be evaluated with the `pinvgamma()` function:
```{r}
f <- function(x) dchi(x, 7)
q <- 2
integrate(f, 0, q)
(p <- pchi(q, 7))
```

The [quantile function](https://en.wikipedia.org/wiki/Quantile_function) can be evaluated with `qchi()`:
```{r}
qchi(p, 7) # = q
```

And random number generation can be performed with `rchi()`:
```{r}
set.seed(1)
rchi(5, 7)
```
`rchi()` can be used to obtain a [Monte Carlo](https://en.wikipedia.org/wiki/Monte_Carlo_method) estimate of the probability given by `pchi()` above:
```{r}
samples <- rchi(1e5, 7)
mean(samples <= q)
```
Moreover, we can check the consistency and correctness of the implementation with
```{r, fig.height = 3}
qplot(samples, geom = "density") + 
  stat_function(fun = f,  color = "red")
```




### Related packages

The __Runuran__ package also implements a sampler from the chi distribution, using the function `Runuran::urchi()`.  (It also provides `udchi()`, but not `upchi()` or `uqchi()`.)  Here's a comparison of how fast the samplers are, with the punch line being that __Runuran__ is about 3x faster for larger numbers of samples, whereas __chi__ is much faster for smaller samples (e.g. 100 or 1000).

```{r}
library(Runuran)
library(microbenchmark)

# chi::rchi() is much faster for small datasets
microbenchmark(
  urchi(1e3, 5),
  rchi(1e3, 5)
)

# Runuran::urchi is ~3x faster for larger datasets
microbenchmark(
  urchi(1e5, 5),
  rchi(1e5, 5)
)
```

























